### 2024-03-25 Introduction+SummaryStats.Rmd
Data set is downloaded, cleaned, and initial statistics are done. Considerations for future: there are some missing weeks; how do we make sure to include that there are skips in the data?

```{r}
library(ggplot2)
library(readxl)
library(cowplot)
library(tidyverse)
library(lubridate)
library(dplyr)
```

```{r}
TempData <- read_excel("Temperature2011-2021 (1).xlsx")
TempData <- TempData |>
  mutate(
    `Decimal Years` = date(date_decimal(`Decimal Years`))) |>
  na.omit()

  colnames(TempData) <- c("Date", "Temp", "SD")
```

```{r}
head(TempData)
```

```{r}
ggplot(TempData)+
  geom_point(aes(x = Date, y= Temp)) +
  labs(x = "Year", y = "Temperature") 
```

```{r}
Summary <- TempData |>
  group_by(year(Date)) |>
  summarise(Temp = mean(Temp))
```

```{r}
print(Summary)
```

```{r}
ggplot(Summary)+
  geom_point(aes(x = `year(Date)`, y= Temp)) +
  labs(x = "Year", y = "Average Temperature") 
```

```{r}
SeasonTempData <- TempData |>
  mutate(
    Season = case_when(
      month(Date) == 12 | month(Date) == 1 | month(Date) == 2 ~ "Winter",
      month(Date) == 3 | month(Date) == 4 | month(Date) == 5 ~ "Spring",
      month(Date) == 6 | month(Date) == 7 | month(Date) == 8 ~ "Summer",
      month(Date) == 9 | month(Date) == 10 | month(Date) == 11 ~ "Fall"
    )
  )
```

```{r}
ggplot(SeasonTempData)+
  geom_point(aes(x = Date, y= Temp, color = Season)) +
  labs(x = "Year", y = "Temperature") 
```

```{r}
SummarySeason <- SeasonTempData |>
  group_by(year(Date), Season) |>
  summarise(Temp = mean(Temp))
```

```{r}
print(SummarySeason)
```

```{r}
SummarySeasonSummer <- SummarySeason |>
  filter(Season == "Summer")
print(SummarySeasonSummer)
```

```{r}
SummarySeasonSummer <- SummarySeason |>
  filter(Season == "Winter")
print(SummarySeasonSummer)
```

```{r}
SummarySeasonSummer <- SummarySeason |>
  filter(Season == "Spring")
print(SummarySeasonSummer)
```

```{r}
SummarySeasonSummer <- SummarySeason |>
  filter(Season == "Fall")
print(SummarySeasonSummer)
```

### 2024-03-25 Introduction+SummaryStats.Rmd
fixed names

@@
 library(dplyr)
+library(knitr)
 ```
@@
 ```{r}
-SummarySeasonSummer <- SummarySeason |>
+SummarySeasonWinter <- SummarySeason |>
   filter(Season == "Winter")
@@
 ```{r}
-SummarySeasonSummer <- SummarySeason |>
+SummarySeasonSpring <- SummarySeason |>
   filter(Season == "Spring")
@@
 ```{r}
-SummarySeasonSummer <- SummarySeason |>
+SummarySeasonFall <- SummarySeason |>
   filter(Season == "Fall")

### 2024-04-09 Introduction+SummaryStats.Rmd
new data set added, NAs are not removed (need to replace with averages)

@@
 ```{r}
-TempData <- read_excel("Temperature2011-2021 (1).xlsx")
+TempData <- read_excel("~/WaterTemperature_TSA_FinalProject/picodataexport-temperatureNew.xlsx", col_names = FALSE,     col_types = c("numeric", "numeric", "numeric"))
+
+colnames(TempData) <- c("Date", "Temp", "SD")
+
 TempData <- TempData |>
   mutate(
-    `Decimal Years` = date(date_decimal(`Decimal Years`))) |>
-  na.omit()
+    Date = date(date_decimal(Date)))
 
-  colnames(TempData) <- c("Date", "Temp", "SD")
 ```

### 2024-04-15 Draft_report.Rmd

---
title: "Final Project"
subtitle: "Water Temperature"
author: "Zhanylai Turatkhan kyzy, Julia Kagiliery, Yilun Zhu"
output: pdf_document
geometry: margin=2.54cm
---

**Introduction**

**Motivation/relevance of the study**

**Objectives**

**Dataset Information**

**Methodology / Analysis**

**Description**
1. SARIMA model
2. TBATS model
3. ETS model
4. Neural Network
5. SSES model

```{r, echo=FALSE}
library(ggplot2)
library(readxl)
library(cowplot)
library(tidyverse)
library(lubridate)
library(dplyr)
library(knitr)
library(caret)
library(forecast)
```

```{r, echo=FALSE}
TempData <- read_excel("Temperature2011-2021 (1).xlsx")
TempData <- TempData |>
  mutate(
    `Decimal Years` = date(date_decimal(`Decimal Years`))) |>
  na.omit()

  colnames(TempData) <- c("Date", "Temp", "SD")
```

```{r, echo=FALSE}
head(TempData)
```

```{r, echo=FALSE}
set.seed(123)

index <- createDataPartition(TempData$Temp, p=0.8, list=FALSE)
trainingData <- TempData[index, ]
testData <- TempData[-index, ]
```

```{r, echo=FALSE}
ggplot(trainingData, aes(x=Date, y=Temp)) +
  geom_point() +
  theme_minimal() +
  labs(title = "Temperature over time",
       x = "Date",
       y = "Temperature (°C)")
```
The plot illustrates the temperature variations over time. Here we can see potential seasonal patterns in the vertical alignment of points across the years.

```{r, echo=FALSE}
par(mar=c(3,3,3,0));par(mfrow=c(1,2))
ACF_plot <- Acf(trainingData$Temp, lag=104, main="Autocorrelation of Temperature")
PACF_plot <- Pacf(trainingData$Temp, lag=104, main="Partial autocorr. of Temperature")
```

The clear seasonal pattern in the ACF plot suggests that any time series 
forecasting models should account for seasonal effects, likely including 
seasonal differencing at lag 52 to remove the seasonal component.
The PACF plot suggests considering a few autoregressive terms in the model, 
as indicated by the initial significant lags. The insignificance of the PACF 
after the first few lags suggests that a simple AR model may be sufficient for 
the non-seasonal part of the model, whereas the significant ACF at seasonal 
lags points to the need for a seasonal AR component.

```{r, echo=FALSE}
TempData <- TempData[order(TempData$Date),]

temperature_ts <- ts(TempData$Temp, frequency = 52, 
                     start = c(year(TempData$Date[1]), 
                               week(TempData$Date[1])))
```

```{r, echo=FALSE}}
start(temperature_ts)
end(temperature_ts)
```

### 2024-04-15 Draft_report.Rmd
fixed time series full dataset and split into train_set and data_set.

@@
 3. ETS model
-4. Neural Network
+4. 
 5. SSES model
@@
 library(forecast)
+library(zoo)
 ```
@@
 lags points to the need for a seasonal AR component.
+```{r, echo=FALSE}
+print(length(TempData$Temp))
+```
+```{r, echo=FALSE}
+# Assuming TempData has columns Date and Temp
+TempData$Date <- as.Date(TempData$Date)
+zoo_ts <- zoo(TempData$Temp, order.by = TempData$Date)
+
+# Check the structure
+print(zoo_ts)
+```
 
 ```{r, echo=FALSE}
-TempData <- TempData[order(TempData$Date),]
-
-temperature_ts <- ts(TempData$Temp, frequency = 52, 
-                     start = c(year(TempData$Date[1]), 
-                               week(TempData$Date[1])))
+start(zoo_ts)
+end(zoo_ts)
+```
+```{r, echo=FALSE}
+split_index <- floor(0.8 * length(zoo_ts))
+train_set <- zoo_ts[1:split_index]
+test_set <- zoo_ts[(split_index + 1):length(zoo_ts)]
 ```
 
-```{r, echo=FALSE}}
-start(temperature_ts)
-end(temperature_ts)
+```{r, echo=FALSE}
+print(head(train_set))
+print(head(test_set))
 ```
+
+```{r, echo=FALSE}
+start(train_set)
+end(train_set)
+```
+
+```{r, echo=FALSE}
+start(test_set)
+end(test_set)
+```
+
+```{r, echo=FALSE}
+fit <- auto.arima(temperature_ts, seasonal = TRUE)
+```
+
+```{r, echo=FALSE}
+print(fit)
+```

### 2024-04-16 Draft_report.Rmd
Monthly data and monthly ts object

--- 2024-04-15 Draft_report.Rmd
@@
 3. ETS model
-4. Neural Network
 5. SSES model
+6. Naive
 
@@
 ```{r, echo=FALSE}
-TempData <- read_excel("Temperature2011-2021 (1).xlsx")
-TempData <- TempData |>
-  mutate(
-    `Decimal Years` = date(date_decimal(`Decimal Years`))) |>
-  na.omit()
-
-  colnames(TempData) <- c("Date", "Temp", "SD")
+file_path <- "/home/guest/TSA_Janka/WaterTemperature_TSA_FinalProject/Monthly Data_Temperature2011-2021.xlsx"
+temperature_data <- read_excel(file_path, sheet = "Sheet1")
+head(temperature_data)
+```
+```{r, echo=FALSE}
+temperature_ts <- ts(temperature_data$Temperature,
+                     start = c(2011, 1), frequency = 12)
 ```
@@
 ```{r, echo=FALSE}
-head(TempData)
+plot(temperature_ts)
 ```
-
-```{r, echo=FALSE}
-set.seed(123)
-
-index <- createDataPartition(TempData$Temp, p=0.8, list=FALSE)
-trainingData <- TempData[index, ]
-testData <- TempData[-index, ]
-```
-
-```{r, echo=FALSE}
-ggplot(trainingData, aes(x=Date, y=Temp)) +
-  geom_point() +
-  theme_minimal() +
-  labs(title = "Temperature over time",
-       x = "Date",
-       y = "Temperature (°C)")
-```
-The plot illustrates the temperature variations over time. Here we can see potential seasonal patterns in the vertical alignment of points across the years.
-
-```{r, echo=FALSE}
-par(mar=c(3,3,3,0));par(mfrow=c(1,2))
-ACF_plot <- Acf(trainingData$Temp, lag=104, main="Autocorrelation of Temperature")
-PACF_plot <- Pacf(trainingData$Temp, lag=104, main="Partial autocorr. of Temperature")
-```
-
-The clear seasonal pattern in the ACF plot suggests that any time series 
-forecasting models should account for seasonal effects, likely including 
-seasonal differencing at lag 52 to remove the seasonal component.
-The PACF plot suggests considering a few autoregressive terms in the model, 
-as indicated by the initial significant lags. The insignificance of the PACF 
-after the first few lags suggests that a simple AR model may be sufficient for 
-the non-seasonal part of the model, whereas the significant ACF at seasonal 
-lags points to the need for a seasonal AR component.
-
-```{r, echo=FALSE}
-TempData <- TempData[order(TempData$Date),]
-
-temperature_ts <- ts(TempData$Temp, frequency = 52, 
-                     start = c(year(TempData$Date[1]), 
-                               week(TempData$Date[1])))
-```
-
-```{r, echo=FALSE}}
-start(temperature_ts)
-end(temperature_ts)
-```

### 2024-04-16 Yilun Upload/YilunCode.Rmd
Yilun new RMD

---
title: "YilunZhuCode"
output: pdf_document
date: "2024-04-15"
always_allow_html: true
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE) 
```

```{r setup, include=FALSE}
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)
library(tidyverse)
library(smooth)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
Monthly_Data <- read_excel("Monthly Data_Temperature2011-2021.xlsx")
#Transfer Date to only Y/M/D
Monthly_Data$Date <- format(as.Date(Monthly_Data$Date), "%Y/%m")

```

```{r}
ts_Temperature_Monthly <- msts(Monthly_Data$Temperature, 
                           seasonal.periods =c(12),
                           start=c(2011,1)
                           )
#create a subset for training purpose
n_for = 24
ts_Temperature_Monthly_train <- subset(ts_Temperature_Monthly,
                                   end = length(ts_Temperature_Monthly)-n_for)

#create a subset for testing purpose
ts_Temperature_Monthly_test <- subset(ts_Temperature_Monthly,
                                   start = length(ts_Temperature_Monthly)-n_for)
```

```{r}
#NN_fit <- nnetar(ts_act_power_daily_train,p=1,P=1)
NN_fit <- nnetar(ts_Temperature_Monthly_train,
                 p=1,
                 P=1,
                 xreg=fourier(ts_Temperature_Monthly_train, K=c(6)))

#NN_for <- forecast(NN_fit, h=6) 
NN_for <- forecast(NN_fit, h=24,xreg=fourier(ts_Temperature_Monthly_train, 
                                          K=c(6),h=24))

#Plot foresting results
autoplot(NN_for) +
  ylab("Temperature C")

#Plot model + observed data
autoplot(ts_Temperature_Monthly_test) +
  autolayer(NN_for, series="Neural Network",PI=FALSE)+
  ylab("Temperature")

autoplot(ts_Temperature_Monthly) +
  autolayer(NN_for$fitted, series="NN fit",PI=FALSE) +
  autolayer(NN_for$mean, series="NN forecast",PI=FALSE)+
  ylab("Temperature")

checkresiduals(NN_fit)
```

```{r, echo=FALSE}
SSES_seas <- es(ts_Temperature_Monthly_train,model="ZZZ", h=24,holdout=FALSE)
plot(SSES_seas)
checkresiduals(SSES_seas)

#Plot model + observed data
autoplot(ts_Temperature_Monthly) +
  autolayer(SSES_seas$fitted, series="SSES Fit",PI=FALSE)+
  autolayer(SSES_seas$forecast, series="SSES Forecast",PI=FALSE)+
  ylab("Temperature") 
```

```{r TBATS, echo=TRUE, message=FALSE, warning=FALSE}
# TBATS can take time to fit
TBATS_fit <- tbats(ts_Temperature_Monthly_train)

TBATS_for <- forecast(TBATS_fit, h=24)

#Plot foresting results
autoplot(TBATS_for) +
  ylab("Temperature")

#Plot test model + observed data
autoplot(ts_Temperature_Monthly_test) +
  autolayer(TBATS_for, series="TBATS Forecast",PI=FALSE)+
  ylab("Temperature")

#Plot model + observed data
autoplot(ts_Temperature_Monthly) +
  autolayer(TBATS_fit$fitted.values, series="TBATS Fit",PI=FALSE)+
  autolayer(TBATS_for, series="TBATS Forecast",PI=FALSE)+
  ylab("Temperature")

checkresiduals(TBATS_fit)
```
```{r ETS, echo=TRUE, message=FALSE, warning=FALSE}
#Fit and forecast STL + ETS model to data
ETS_fit <-  stlf(ts_Temperature_Monthly_train,h=24)
#Since stlf did both model+forcastm so h =365 is required to specify the edge of forcast

#Plot foresting results
autoplot(ETS_fit) + ylab("Temperature")

#Plot model + observed data
autoplot(ts_Temperature_Monthly) +
  autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
  ylab("Temperature")

#Plot model + observed data
autoplot(ts_Temperature_Monthly) +
  autolayer(ETS_fit$fitted, series="STL + ETS training",PI=FALSE) +
  autolayer(ETS_fit$mean, series="STL + ETS test",PI=FALSE) +
  ylab("Temperature")

checkresiduals(ETS_fit)
#alternative coding if you don't need to save the objects
#ts_act_power_daily_train %>% stlf(h=365) %>% autoplot() 
#very easy to read/run/implement 
#but you don't have anything stored on your environment

```

```{r ARIMA, echo=TRUE, message=FALSE, warning=FALSE}
#Fit arima model with fourier terms as exogenous regressors
# seasonal = FALSE is the same as P=D=Q=0
# play with K by changing it to K=c(2,2), K=c(2,4), K=c(2,6), etc. The higher the K the longer it will take to converge, because R will try more models.

ARIMA_Four_fit <- auto.arima(ts_Temperature_Monthly_train, 
                             seasonal=TRUE, 
                             lambda=0,
                             xreg=fourier(ts_Temperature_Monthly_train, 
                                          K=c(6))
                             )

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Four_for <- forecast(ARIMA_Four_fit,
                           xreg=fourier(ts_Temperature_Monthly_train,
                                        K=c(6),
                                        h=24),
                           h=24
                           )

#Plot foresting results
autoplot(ARIMA_Four_for) + ylab("Temperature")

#Plot model + observed data
autoplot(ts_Temperature_Monthly) +
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE) +
  ylab("Temperature")

```
```{r}

#Model 1: STL + ETS
ETS_scores <- accuracy(ETS_fit$mean,ts_Temperature_Monthly_test)

#Model 2: ARIMA + Fourier 
ARIMA_scores <- accuracy(ARIMA_Four_for$mean,ts_Temperature_Monthly_test)

# Model 3:  TBATS 
TBATS_scores <- accuracy(TBATS_for$mean,ts_Temperature_Monthly_test)

# Model 4:  Neural Network 
NN_scores <- accuracy(NN_for$mean,ts_Temperature_Monthly_test)

# Model 5:  SSES 
SSES_scores <- accuracy(SSES_seas$forecast,ts_Temperature_Monthly_test)

```

```{r}
#create data frame
scores <- as.data.frame(
  rbind(ETS_scores, ARIMA_scores, TBATS_scores, NN_scores)
  )
row.names(scores) <- c("STL+ETS", "ARIMA+Fourier","TBATS","NN","SSES")

#choose model with lowest RMSE
best_model_index <- which.min(scores[,"RMSE"])
cat("The best model by RMSE is:", row.names(scores[best_model_index,]))                       
```

```{r}
autoplot(ts_Temperature_Monthly) +
  autolayer(ETS_fit, PI=FALSE, series="STL+ETS") +
  autolayer(ARIMA_Four_for, PI=FALSE, series="ARIMA + Fourier") +
  autolayer(TBATS_for,PI=FALSE, series="TBATS") +
  autolayer(NN_for,PI=FALSE, series="NN") +
  xlab("Monthly") + ylab("Water Temperature") +
  guides(colour=guide_legend(title="Forecast"))

```

### 2024-04-16 Draft_report.Rmd

@@
 3. ETS model
-5. SSES model
-6. Naive
+4. SSES model
+5. NN
 
@@
 ```
+```{r}
+str(temperature_ts)
+```
+```{r}
+split_point <- round(length(temperature_ts) * 0.8)
+train_ts <- window(temperature_ts, end=c(2019, 8))
+test_ts <- window(temperature_ts, start=c(2019, 9))
+```
+```{r}
+start(train_ts)
+end(train_ts)
+```
+```{r}
+start(test_ts)
+end(test_ts)
+```
+
+```{r, echo=FALSE}
+#SARIMA autofit
+sarima_model <- auto.arima(train_ts)
+print(summary(sarima_model))
+```
+```{r}
+checkresiduals(sarima_model)
+```
+
+```{r}
+h <- length(test_ts)
+sarima_forecast <- forecast(sarima_model, h=h)
+```
+```{r}
+autoplot(temperature_ts, series = "Original") +
+  autolayer(sarima_forecast$mean, series = "SARIMA forecast") +
+  ylab("Water Temperature") +
+  ggtitle("SARIMA modeling")
+```
+
+```{r}
+checkresiduals(sarima_forecast)
+```
+```{r}
+forecast_accuracy <- accuracy(sarima_forecast, test_ts)
+print(forecast_accuracy)
+```
+```{r}
+# STL + ETS
+ets_model <- stlf(train_ts, h=h)
+```
+```{r}
+autoplot(temperature_ts, series = "Original") +
+  autolayer(ets_model, series = "STL + ETS", PI=FALSE) +
+  ylab("Water Temperature") +
+  ggtitle("STL + ETS modeling")
+```
+
+```{r}
+stlf_accuracy <- accuracy(ets_model, test_ts)
+print(stlf_accuracy)
+```
+```{r}
+checkresiduals(ets_model)
+```

### 2024-04-16 Draft_report.Rmd
added ACF, PACF, and Tbats model. Reformat and added code comments

--- 2024-04-16 Draft_report.Rmd
@@
 ---
-
-**Introduction**
-
-**Motivation/relevance of the study**
-
-**Objectives**
-
-**Dataset Information**
-
-**Methodology / Analysis**
-
-**Description**
-1. SARIMA model
-2. TBATS model
-3. ETS model
-5. SSES model
-6. Naive
 
@@
 
+**Introduction**
+
+**Motivation/relevance of the study**
+
+**Objectives**
+
+**Dataset Information**
+
 ```{r, echo=FALSE}
-file_path <- "/home/guest/TSA_Janka/WaterTemperature_TSA_FinalProject/Monthly Data_Temperature2011-2021.xlsx"
+#We important the monthly temperature data
+file_path <- "~/WaterTemperature_TSA_FinalProject/Monthly Data_Temperature2011-2021.xlsx"
 temperature_data <- read_excel(file_path, sheet = "Sheet1")
-head(temperature_data)
 ```
+
 ```{r, echo=FALSE}
+#Transforming the data into a time series which starts in Jan of 2011
 temperature_ts <- ts(temperature_data$Temperature,
@@
 ```{r, echo=FALSE}
+#Upon initial examination, the data appears to be strongly seasonal with very little trend.
 plot(temperature_ts)
 ```
+
+```{r}
+#The ACF and Pacf tell us that there is ovbious and strong seasonality and some auto-regressive component up to about 6 lags.
+P1 <-Acf(temperature_ts)
+P2 <- Pacf(temperature_ts)
+```
+
+HERE WE SHOULD DECOMPOSE THE DATA.
+
+```{r}
+#confirming that timeseries transformation works as expected.
+str(temperature_ts)
+```
+
+In the lines of code below we split the existing data set into a training data set and a test set. 80% of the data is training and the last 20% is used to validate the model.
+```{r}
+#Making Training and Test sets
+split_point <- round(length(temperature_ts) * 0.8)
+train_ts <- window(temperature_ts, end=c(2019, 8))
+test_ts <- window(temperature_ts, start=c(2019, 9))
+```
+
+```{r}
+#Confirms Training and Test sets were correctly made
+start(train_ts)
+end(train_ts)
+start(test_ts)
+end(test_ts)
+```
+**Methodology / Analysis**
+
+**Description**
+1. SARIMA model
+2. TBATS model
+3. ETS model
+4. SSES model
+5. NN
+
+*SARIMA Modeling*
+```{r, echo=FALSE}
+#SARIMA autofit
+sarima_model <- auto.arima(train_ts)
+print(summary(sarima_model))
+```
+
+```{r}
+checkresiduals(sarima_model)
+```
+
+```{r}
+h <- length(test_ts)
+sarima_forecast <- forecast(sarima_model, h=h)
+```
+
+```{r}
+autoplot(temperature_ts, series = "Original") +
+  autolayer(sarima_forecast$mean, series = "SARIMA forecast") +
+  ylab("Water Temperature") +
+  ggtitle("SARIMA modeling")
+```
+
+```{r}
+checkresiduals(sarima_forecast)
+```
+
+```{r}
+forecast_accuracySarima <- accuracy(sarima_forecast, test_ts)
+print(forecast_accuracySarima)
+```
+
+*STL + ETS Modeling*
+
+```{r}
+# STL + ETS
+ets_model <- stlf(train_ts, h=h)
+```
+
+```{r}
+autoplot(temperature_ts, series = "Original") +
+  autolayer(ets_model, series = "STL + ETS", PI=FALSE) +
+  ylab("Water Temperature") +
+  ggtitle("STL + ETS modeling")
+```
+
+```{r}
+stlf_accuracy <- accuracy(ets_model, test_ts)
+print(stlf_accuracy)
+```
+
+```{r}
+checkresiduals(ets_model)
+```
+
+*TBATS Modeling*
+
+```{r}
+model_tbats <- tbats(train_ts)
+print(summary(model_tbats))
+```
+
+```{r}
+checkresiduals(model_tbats)
+```
+
+```{r}
+TBATS_for <- forecast(model_tbats, h=h)
+
+autoplot(train_ts) +
+  autolayer(TBATS_for, series = "TBATS Forecast", PI = FALSE) +
+  autolayer(test_ts, series = "Test Data", PI = FALSE) +
+  autolayer(fitted(model_tbats), series = "Model", PI = FALSE) +
+  ylab("Water Temperature") +
+  ggtitle("TBATS Modeling")
+
+```
+```{r}
+TBATS_accuracy <- accuracy(TBATS_for, test_ts)
+print(TBATS_accuracy)
+```

### 2024-04-16 Draft_report.Rmd
added relevance and data set intrdouction, fixed some plots

@@
 **Motivation/relevance of the study**
+Global ocean temperature is on the rise which has significant implications for this dynamic and productive biome.
+Among the many services the ocean provides, primary production (of oxygen) is among the most impoartant.
+In fact, approximately 50% of primary production of oxygen comes from marine phytoplankton.
+It has been proven (largely through work by Duke University Marine Lab's Dr.
+Zacakry Johnson and Dr.
+Dana Hunt) that these microbial systems change significantly with seasonal changes (which include dominatntly temperature changes, but also insolation and day length changes).
+These changes alter the way that the ocean cycles nutrients, stores carbon, and produces oxygen.
+Hence, temperature is an important variable in the consideration of much ocean modeling.
+This highlights the imporatnce of acurate temperature prediction.
+Undertsanding what future climate and temperature looks like allows for better prediction of what other ocean cycles will look like.
 
@@
 
+Our objective is to accurately model the monthly temperature of Piver's Island Coastal Observatory and produce reasonable forecasting at appropriate time scales.
+
 **Dataset Information**
+The following data comes from a long running time series study out at Piver's Island Coastal Observatory (PICO) which is located at the Duke University Marine Lab.
+The sytudy monitors the ambient conditions such as turbidity, pH, temperature, and salinity.
+For this study, only temperature was able to be included.
+The temperature is reported as mean monthly temperature.
 
@@
 
+There are first a few important considerations as we get into modeling.
+The first is that this data is actually measured approximately weekly, meaning sometimes the lab samples multiple times in a week or skips a week.
+In order to account for this irregular time series frequency, we aggregated the data into monthly means to avoid unneccesary complications and misalign ment of our time series.
+The second consideration we would like to acknowledge that climate change is certainly a a factor that plays a role in shifting trends, seasonal components, and general variability which may not be accurately reflected in our current data set which spans only approximately 10 years.
+Though the temperature in the region looks highly predicatble, future predictions must be cognizant that large prediction horizions are unreasonable.
+
 **Description**
@@
 ```{r}
-autoplot(temperature_ts, series = "Original") +
-  autolayer(ets_model, series = "STL + ETS", PI=FALSE) +
+ETS_for <- forecast(ets_model, h=h)
+
+autoplot(train_ts) +
+    autolayer(ETS_for, series = "ETS Forecast", PI = FALSE) +
+  autolayer(fitted(ets_model), series = "Model STL + ETS", PI=FALSE) +
+    autolayer(test_ts, series = "Test Data", PI=FALSE) +
   ylab("Water Temperature") +
   ggtitle("STL + ETS modeling")
+
 ```
@@
 ```{r}
-stlf_accuracy <- accuracy(ets_model, test_ts)
+stlf_accuracy <- accuracy(ETS_for, test_ts)
 print(stlf_accuracy)

### 2024-04-16 Yilun Upload/YilunCode.Rmd
Update for MKtest and ADF test

@@
                            )
+decomposed_Temperature <- decompose(ts_Temperature_Monthly, type = "additive")
+plot(decomposed_Temperature)
 #create a subset for training purpose
@@
                                    start = length(ts_Temperature_Monthly)-n_for)
+
+MKTest <- MannKendall(ts_Temperature_Monthly)
+print(summary(MKTest))
+
+print(adf.test(ts_Temperature_Monthly, alternative = "stationary"))
+
 ```

### 2024-04-17 Draft_report.Rmd
Yilun upload code for decompose, MK test, ADF test SSES,NN, ArimaFourier, and socres

@@
 library(zoo)
+library(Kendall)
+library(tseries)
+library(outliers)
+library(smooth)
 ```
@@
 HERE WE SHOULD DECOMPOSE THE DATA.
+```{r}
+decomposed_Temperature <- decompose(temperature_ts, type = "additive")
+plot(decomposed_Temperature)
+
+#Testing if the trend is stationary
+MKTest <- MannKendall(temperature_ts)
+print(summary(MKTest))
+
+print(adf.test(temperature_ts, alternative = "stationary"))
+```
 
@@
 1. SARIMA model
-2. TBATS model
-3. ETS model
-4. SSES model
-5. NN
+2. Arima+Fourier
+3. TBATS model
+4. ETS model
+5. SSES model
+6. Neural Network
 
@@
 ```{r}
-forecast_accuracySarima <- accuracy(sarima_forecast, test_ts)
+forecast_accuracySarima <- accuracy(sarima_forecast)
 print(forecast_accuracySarima)
+```
+
+*ARIMA+fourier Modeling*
+```{r ARIMA, echo=TRUE, message=FALSE, warning=FALSE}
+ARIMA_Four_fit <- auto.arima(train_ts, 
+                             seasonal=TRUE, 
+                             lambda=0,
+                             xreg=fourier(train_ts, 
+                                          K=c(6))
+                             )
+
+ARIMA_Four_for <- forecast(ARIMA_Four_fit,
+                           xreg=fourier(train_ts,
+                                        K=c(6),
+                                        h=h),
+                           h=h
+                           )
+
+autoplot(ARIMA_Four_for) + ylab("Temperature")
+
+autoplot(temperature_ts) +
+  autolayer(ARIMA_Four_fit$fitted, series="ARIMA_FOURIER Fitted",PI=FALSE) +
+  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER Forecast",PI=FALSE) +
+  ylab("Temperature")
+
+checkresiduals(ARIMA_Four_for)
+
+forecast_accuracyARIMA_Four <- accuracy(ARIMA_Four_for)
+print(forecast_accuracyARIMA_Four)
 ```
@@
 ```{r}
-stlf_accuracy <- accuracy(ETS_for, test_ts)
+stlf_accuracy <- accuracy(ETS_for)
 print(stlf_accuracy)
@@
 ```{r}
-TBATS_accuracy <- accuracy(TBATS_for, test_ts)
+TBATS_accuracy <- accuracy(TBATS_for)
 print(TBATS_accuracy)
 ```
+
+*SSES*
+```{r, echo=FALSE}
+SSES_seas <- es(train_ts,model="ZZZ", h= h,holdout=FALSE)
+plot(SSES_seas)
+checkresiduals(SSES_seas)
+
+#Plot model + observed data
+autoplot(temperature_ts) +
+  autolayer(SSES_seas$fitted, series="SSES Fit",PI=FALSE)+
+  autolayer(SSES_seas$forecast, series="SSES Forecast",PI=FALSE)+
+  ylab("Temperature")
+
+forecast_accuracy_SSES <- accuracy(SSES_seas)
+```
+
+*Neural Network*
+```{r}
+NN_fit <- nnetar(train_ts,
+                 p=1,
+                 P=1,
+                 xreg=fourier(train_ts, K=c(6)))
+
+NN_for <- forecast(NN_fit, h= h,xreg=fourier(train_ts, 
+                                          K=c(6),h= h))
+
+autoplot(NN_for) +
+  ylab("Temperature C")
+
+autoplot(test_ts) +
+  autolayer(NN_for, series="Neural Network",PI=FALSE)+
+  ylab("Temperature")
+
+autoplot(temperature_ts) +
+  autolayer(NN_for$fitted, series="NN fit",PI=FALSE) +
+  autolayer(NN_for$mean, series="NN forecast",PI=FALSE)+
+  ylab("Temperature")
+
+checkresiduals(NN_fit)
+
+forecast_accuracy_NN <- accuracy(NN_for)
+```
+*Create Scores*
+```{r}
+
+#Model 1: Sarima
+Sarima_scores <- accuracy(sarima_forecast$mean, test_ts)
+
+#Model 2: ARIMA + Fourier
+ArimaFour_scores <- accuracy(ARIMA_Four_for$mean, test_ts)
+
+#Model 3: ETS+STL 
+stlf_scores <- accuracy(ETS_for$mean, test_ts)
+
+# Model 4:  TBATS 
+TBATS_scores <- accuracy(TBATS_for$mean, test_ts)
+
+# Model 5:  SSES 
+SSES_scores <- accuracy(NN_for$mean, test_ts)
+
+# Model 6:  Neural Network 
+NN_scores <- accuracy(SSES_seas$forecast, test_ts)
+
+```
+
+*Create Score table*
+```{r}
+scores <- as.data.frame(
+  rbind(Sarima_scores, ArimaFour_scores, stlf_scores, TBATS_scores, SSES_scores, NN_scores)
+  )
+row.names(scores) <- c("Sarima", "ARIMA+Fourier","ETS+STL","TBATS","SSES", "NN")
+
+best_model_index <- which.min(scores[,"RMSE"])
+cat("The best model by RMSE is:", row.names(scores[best_model_index,]))                       
+```

### 2024-04-17 Yilun Upload/YilunCode.Rmd
Yilun upload code for decompose, MK test, ADF test SSES,NN, ArimaFourier, and socres

@@
                                    start = length(ts_Temperature_Monthly)-n_for)
-
+#Testing if the trend is stationary
 MKTest <- MannKendall(ts_Temperature_Monthly)

### 2024-04-22 Draft_report.Rmd
90% of report done

@@
 **Motivation/relevance of the study**
-Global ocean temperature is on the rise which has significant implications for this dynamic and productive biome.
-Among the many services the ocean provides, primary production (of oxygen) is among the most impoartant.
-In fact, approximately 50% of primary production of oxygen comes from marine phytoplankton.
-It has been proven (largely through work by Duke University Marine Lab's Dr.
-Zacakry Johnson and Dr.
-Dana Hunt) that these microbial systems change significantly with seasonal changes (which include dominatntly temperature changes, but also insolation and day length changes).
-These changes alter the way that the ocean cycles nutrients, stores carbon, and produces oxygen.
-Hence, temperature is an important variable in the consideration of much ocean modeling.
-This highlights the imporatnce of acurate temperature prediction.
-Undertsanding what future climate and temperature looks like allows for better prediction of what other ocean cycles will look like.
+Global ocean temperature is on the rise which has significant implications for 
+this dynamic and productive biome. Among the many services the ocean provides, 
+primary production (of oxygen) is among the most important. In fact, 
+approximately 50% of primary production of oxygen comes from marine 
+phytoplankton. It has been proven (largely through work by Duke University 
+Marine Lab's Dr. Zacakry Johnson and Dr. Dana Hunt) that these microbial 
+systems change significantly with seasonal changes (which include dominantly temperature changes, but also isolation and day length changes). These changes 
+alter the way that the ocean cycles nutrients, stores carbon, and produces 
+oxygen. Hence, temperature is an important variable in the consideration of 
+much ocean modeling. This highlights the importance of accurate temperature 
+prediction. Understanding what future climate and temperature looks like 
+allows for better prediction of what other ocean cycles will look like.
 
@@
 
-Our objective is to accurately model the monthly temperature of Piver's Island Coastal Observatory and produce reasonable forecasting at appropriate time scales.
+Our objective is to accurately model the monthly temperature of Piver's 
+Island Coastal Observatory and produce reasonable forecasting at appropriate 
+time scales.
 
 **Dataset Information**
-The following data comes from a long running time series study out at Piver's Island Coastal Observatory (PICO) which is located at the Duke University Marine Lab.
-The sytudy monitors the ambient conditions such as turbidity, pH, temperature, and salinity.
-For this study, only temperature was able to be included.
-The temperature is reported as mean monthly temperature.
+
+The following data comes from a long running time series study out at Piver's 
+Island Coastal Observatory (PICO) which is located at the Duke University 
+Marine Lab. The sytudy monitors the ambient conditions such as turbidity, pH, temperature, and salinity. For this study, only temperature was able to be 
+included. The temperature is reported as mean monthly temperature.
+
+**Methodology / Analysis**
+
+There are first a few important considerations as we get into modeling. The 
+first is that this data is actually measured approximately weekly, meaning 
+sometimes the lab samples multiple times in a week or skips a week. In order 
+to account for this irregular time series frequency, we aggregated the data 
+into monthly means to avoid unnecessary complications and misalignment of our 
+time series. 
+The second consideration we would like to acknowledge that climate change is 
+certainly a a factor that plays a role in shifting trends, seasonal components, 
+and general variability which may not be accurately reflected in our current 
+data set which spans only approximately 11 years. Though the temperature in the 
+region looks highly predictable, future predictions must be cognizant that large prediction horizons are unreasonable.
+
+**Description**
+
+**1. SARIMA model:** The Seasonal Autoregressive Integrated Moving Average 
+(SARIMA) model expands upon the ARIMA framework to address seasonality in 
+univariate data sets. It incorporates both non-seasonal and seasonal elements 
+in its predictions, allowing it to capture complex patterns that recur over 
+fixed periods.
+**2. Arima+Fourier:** This approach combines the ARIMA model with Fourier series 
+to enhance the modeling of time series with complex seasonal patterns. 
+ARIMA captures the autocorrelations in the data, while the Fourier terms allow 
+for the approximation of seasonal cycles of various lengths and complexities. 
+This hybrid model is especially useful when the seasonality is not strictly 
+periodic or involves multiple frequencies.
+**3. TBATS model:** Designed for forecasting time series with intricate seasonal patterns, the TBATS model employs exponential smoothing as its core technique. 
+It thoroughly explores a variety of specifications, including those with and 
+without a Box-Cox transformation, the presence or absence of a trend, trend 
+damping options, and an ARIMA(p,q) component for the residuals. The model also evaluates different harmonic levels for seasonalities. The best-fitting model 
+is determined by the lowest Akaike Information Criterion (AIC).
+**4. ETS model:** This univariate forecasting method, known as Exponential 
+Smoothing State Space Model (ETS), emphasizes trend and seasonality components 
+within the time series data. It is particularly adept at capturing patterns 
+that evolve over time.
+**5. SSES model:** The State Space Exponential Smoothing (SSES) model extends 
+the classic exponential smoothing approach by incorporating distribution 
+assumptions about the error terms, which aids in the computation of prediction 
+intervals. It considers both additive and multiplicative error structures 
+within the state space modeling framework.
+**6. Neural Network:** The Neural Network model facilitates the identification 
+of complex and nonlinear relationships between the dependent variable and its predictors. Its versatile architecture can adapt to a wide array of data 
+patterns.
+
+```{r, echo=FALSE}
+#file_path <- file.path(getwd(), "Monthly Data_Temperature2011-2021.xlsx")
+#print(file_path)
+```
 
@@
 #We important the monthly temperature data
-file_path <- "~/WaterTemperature_TSA_FinalProject/Monthly Data_Temperature2011-2021.xlsx"
+file_path <- "/home/guest/TSA_Janka/WaterTemperature_TSA_FinalProject/Monthly Data_Temperature2011-2021.xlsx"
 temperature_data <- read_excel(file_path, sheet = "Sheet1")
@@
 #Upon initial examination, the data appears to be strongly seasonal with very little trend.
-plot(temperature_ts)
+plot(temperature_ts, main = "Original dataset")
 ```
+The temperature shows cyclical behavior with peaks and troughs that correspond 
+to expected seasonal variations. The amplitude of these oscillations appears 
+consistent over the years, indicating a stable seasonal effect without 
+significant year-over-year changes in peak or trough temperatures.
+From a visual inspection, there are no apparent outliers or disruptions in the 
+seasonal pattern, suggesting that the dataset is clean and well-maintained. 
+The regularity of the pattern would likely make it suitable for forecasting 
+using seasonal models, such as SARIMA or TBATS, which could exploit the 
+periodicity inherent in the data.
+```{r, echo=FALSE}
+#The ACF and Pacf tell us that there is ovbious and strong seasonality and some auto-regressive component up to about 6 lags.
+P1 <-Acf(temperature_ts, main = "ACF on original data")
+```
+The graph illustrates the Autocorrelation Function (ACF) applied to the original temperature data set. Autocorrelation is a mathematical representation of the 
+degree of similarity between a given time series and a lagged version of itself 
+over successive time intervals. The ACF is plotted against various lag values, 
+which span from 0 to 24.
+The y-axis represents the autocorrelation coefficient, ranging from -0.5 to 0.5, 
+while the x-axis represents the lag in terms of the number of time units. Each 
+vertical bar corresponds to the autocorrelation coefficient at a specific lag.
+Notably, the graph shows a pattern of spikes at regular intervals, which 
+suggests a seasonal pattern in the data. The presence of these spikes at 
+consistent intervals can be indicative of the seasonality in the dataset, which correlates with the seasonal fluctuations seen in the time series plot of 
+the original dataset.
+The blue dashed lines represent the significance bounds. Any spike that extends 
+beyond these bounds is considered statistically significant. In this ACF plot, 
+several lags have autocorrelation values that cross the significance threshold,
+confirming that the data exhibit a strong seasonal component at these lags.
 
-```{r}
-#The ACF and Pacf tell us that there is ovbious and strong seasonality and some auto-regressive component up to about 6 lags.
-P1 <-Acf(temperature_ts)
-P2 <- Pacf(temperature_ts)
+```{r, echo=FALSE}}
+P2 <- Pacf(temperature_ts, main = "PACF on original data")
 ```
+In this PACF plot, most of the spikes are within the significance bounds, 
+suggesting that most of the autocorrelations in the original data can be 
+accounted for by the immediate preceding values.
 
-HERE WE SHOULD DECOMPOSE THE DATA.
-```{r}
+```{r, echo=FALSE}
 decomposed_Temperature <- decompose(temperature_ts, type = "additive")
@@
 ```
-
-```{r}
+The Augmented Dickey-Fuller Test result indicates a Dickey-Fuller statistic of 
+-11.153 with a lag order of 5, and a p-value of 0.01, suggesting that the null 
+hypothesis of a unit root can be rejected and the time series is stationary.
+```{r, echo=FALSE}
 #confirming that timeseries transformation works as expected.
-str(temperature_ts)
+#str(temperature_ts)
 ```
 
-In the lines of code below we split the existing data set into a training data set and a test set. 80% of the data is training and the last 20% is used to validate the model.
-```{r}
+For the forecasting purposes, the data was split into training and test data
+following a proportion of 80/20.
+```{r, echo=FALSE}
 #Making Training and Test sets
@@
 
-```{r}
+```{r, echo=FALSE}
 #Confirms Training and Test sets were correctly made
-start(train_ts)
-end(train_ts)
-start(test_ts)
-end(test_ts)
+#start(train_ts)
+#end(train_ts)
+#start(test_ts)
+#end(test_ts)
 ```
-**Methodology / Analysis**
-
-There are first a few important considerations as we get into modeling.
-The first is that this data is actually measured approximately weekly, meaning sometimes the lab samples multiple times in a week or skips a week.
-In order to account for this irregular time series frequency, we aggregated the data into monthly means to avoid unneccesary complications and misalign ment of our time series.
-The second consideration we would like to acknowledge that climate change is certainly a a factor that plays a role in shifting trends, seasonal components, and general variability which may not be accurately reflected in our current data set which spans only approximately 10 years.
-Though the temperature in the region looks highly predicatble, future predictions must be cognizant that large prediction horizions are unreasonable.
-
-**Description**
-1. SARIMA model
-2. Arima+Fourier
-3. TBATS model
-4. ETS model
-5. SSES model
-6. Neural Network
 
@@
 
-```{r}
+```{r, echo=FALSE}
 checkresiduals(sarima_model)
@@
 
-```{r}
+```{r, echo=FALSE}
 h <- length(test_ts)
@@
 
-```{r}
+```{r, echo=FALSE}
 autoplot(temperature_ts, series = "Original") +
@@
 
-```{r}
+```{r, echo=FALSE}
 checkresiduals(sarima_forecast)
@@
 
-```{r}
+```{r, echo=FALSE}
 forecast_accuracySarima <- accuracy(sarima_forecast)
@@
 
-```{r}
+```{r, echo=FALSE}
 # STL + ETS
@@
 
-```{r}
+```{r, echo=FALSE}
 ETS_for <- forecast(ets_model, h=h)
@@
 
-```{r}
+```{r, echo=FALSE}
 stlf_accuracy <- accuracy(ETS_for)
@@
 
-```{r}
+```{r, echo=FALSE}
 checkresiduals(ets_model)
@@
 
-```{r}
+```{r, echo=FALSE}
 model_tbats <- tbats(train_ts)
@@
 
-```{r}
+```{r, echo=FALSE}
 checkresiduals(model_tbats)
@@
 
-```{r}
+```{r, echo=FALSE}
 TBATS_for <- forecast(model_tbats, h=h)
@@
 ```
-```{r}
+```{r, echo=FALSE}
 TBATS_accuracy <- accuracy(TBATS_for)
@@
 *Neural Network*
-```{r}
+```{r, echo=FALSE}
 NN_fit <- nnetar(train_ts,
@@
 *Create Scores*
-```{r}
+```{r, echo=FALSE}
 
@@
 *Create Score table*
-```{r}
+```{r, echo=FALSE}
 scores <- as.data.frame(

### 2024-04-22 Draft_report.Rmd
added limiatations, intro and refrneces

@@
 **Introduction**
+Marine microbe account for 50% of marine primary production and dominate the 
+biomass of the ocean. These organism are vital for life on earth and ocean 
+health but are at serious risk in high temperatures worsened by climate change. 
+The first step in supporting these ecological communities and hence mitigating 
+the damages done by climate change is to understand what future climate 
+scenarios look like. Predictions for future climate and temperature can varry by 
+up to 10 degrees (if not more). With so much uncertainty surrounding climate 
+future, it is of the utmost importance to have as much acurate modeling for 
+future climate scenarios and hence our group hopes to make reasonable forecasts 
+of surface water temperature at the Duke University Marine Lab research facility.
 
@@
 Marine Lab's Dr. Zacakry Johnson and Dr. Dana Hunt) that these microbial 
-systems change significantly with seasonal changes (which include dominantly temperature changes, but also isolation and day length changes). These changes 
+systems change significantly with seasonal changes (which include dominantly 
+temperature changes, but also isolation and day length changes). These changes 
 alter the way that the ocean cycles nutrients, stores carbon, and produces 
-oxygen. Hence, temperature is an important variable in the consideration of 
-much ocean modeling. This highlights the importance of accurate temperature 
-prediction. Understanding what future climate and temperature looks like 
-allows for better prediction of what other ocean cycles will look like.
+oxygen. Hence, temperature is an important variable in the consideration of much 
+ocean modeling. This highlights the importance of accurate temperature 
+prediction. Understanding what future climate and temperature looks like allows 
+for better prediction of what other ocean cycles will look like and how we may 
+excpet the serives we recieve from the ocean to change with the climate.
 
@@
 Island Coastal Observatory (PICO) which is located at the Duke University 
-Marine Lab. The sytudy monitors the ambient conditions such as turbidity, pH, temperature, and salinity. For this study, only temperature was able to be 
-included. The temperature is reported as mean monthly temperature.
+Marine Lab conducted by Drs. Zackary Johnson and Dana Hunt. The sytudy monitors 
+the ambient conditions such as turbidity, pH, temperature, and salinity. For 
+this study, only temperature was able to be included. The temperature is 
+reported as mean monthly temperature.
 
@@
 data set which spans only approximately 11 years. Though the temperature in the 
-region looks highly predictable, future predictions must be cognizant that large prediction horizons are unreasonable.
+region looks highly predictable, future predictions must be cognizant that large 
+prediction horizons are unreasonable.
 
@@
 periodic or involves multiple frequencies.
-**3. TBATS model:** Designed for forecasting time series with intricate seasonal patterns, the TBATS model employs exponential smoothing as its core technique. 
+**3. TBATS model:** Designed for forecasting time series with intricate seasonal 
+patterns, the TBATS model employs exponential smoothing as its core technique. 
 It thoroughly explores a variety of specifications, including those with and 
 without a Box-Cox transformation, the presence or absence of a trend, trend 
-damping options, and an ARIMA(p,q) component for the residuals. The model also evaluates different harmonic levels for seasonalities. The best-fitting model 
+damping options, and an ARIMA(p,q) component for the residuals. The model also 
+evaluates different harmonic levels for seasonalities. The best-fitting model 
 is determined by the lowest Akaike Information Criterion (AIC).
@@
 **6. Neural Network:** The Neural Network model facilitates the identification 
-of complex and nonlinear relationships between the dependent variable and its predictors. Its versatile architecture can adapt to a wide array of data 
+of complex and nonlinear relationships between the dependent variable and its 
+predictors. Its versatile architecture can adapt to a wide array of data 
 patterns.
@@
 #We important the monthly temperature data
-file_path <- "/home/guest/TSA_Janka/WaterTemperature_TSA_FinalProject/Monthly Data_Temperature2011-2021.xlsx"
+file_path <-
+  "~/WaterTemperature_TSA_FinalProject/Monthly Data_Temperature2011-2021.xlsx"
 temperature_data <- read_excel(file_path, sheet = "Sheet1")
@@
 ```{r, echo=FALSE}
-#Upon initial examination, the data appears to be strongly seasonal with very little trend.
+#Upon initial examination, the data appears to be seasonal with little trend.
 plot(temperature_ts, main = "Original dataset")
@@
 ```
-The graph illustrates the Autocorrelation Function (ACF) applied to the original temperature data set. Autocorrelation is a mathematical representation of the 
+The graph illustrates the Autocorrelation Function (ACF) applied to the original 
+temperature data set. Autocorrelation is a mathematical representation of the 
 degree of similarity between a given time series and a lagged version of itself 
@@
 suggests a seasonal pattern in the data. The presence of these spikes at 
-consistent intervals can be indicative of the seasonality in the dataset, which correlates with the seasonal fluctuations seen in the time series plot of 
+consistent intervals can be indicative of the seasonality in the dataset, which 
+correlates with the seasonal fluctuations seen in the time series plot of 
 the original dataset.
@@
 
-```{r, echo=FALSE}}
+```{r, echo=FALSE}
 P2 <- Pacf(temperature_ts, main = "PACF on original data")
@@
 #confirming that timeseries transformation works as expected.
-#str(temperature_ts)
+str(temperature_ts)
 ```
@@
 #Confirms Training and Test sets were correctly made
-#start(train_ts)
-#end(train_ts)
-#start(test_ts)
-#end(test_ts)
+start(train_ts)
+end(train_ts)
+start(test_ts)
+end(test_ts)
 ```
@@
 scores <- as.data.frame(
-  rbind(Sarima_scores, ArimaFour_scores, stlf_scores, TBATS_scores, SSES_scores, NN_scores)
+  rbind(
+    Sarima_scores,
+    ArimaFour_scores,
+    stlf_scores,
+    TBATS_scores,
+    SSES_scores,
+    NN_scores
   )
-row.names(scores) <- c("Sarima", "ARIMA+Fourier","ETS+STL","TBATS","SSES", "NN")
+)
+row.names(scores) <-
+  c("Sarima", "ARIMA+Fourier", "ETS+STL", "TBATS", "SSES", "NN")
 
-best_model_index <- which.min(scores[,"RMSE"])
-cat("The best model by RMSE is:", row.names(scores[best_model_index,]))                       
+best_model_index <- which.min(scores[, "RMSE"])
+cat("The best model by RMSE is:", row.names(scores[best_model_index, ]))                       
 ```
+
+**Limitations**
+Though we have produced accurate models and forecasts, there are a few limitations to be aware of.
+The first is regarding our prediction time frame We only have access to ~ 10 years of data which likely does not demonstrate enough of a trend to be recognized as the role of climate change in our models.
+We should expect increasing temepratures in the future as a result of climate change.
+Even if our model did capture this linear trend, the modeling would be based on the assumtions that the trend is constant (aka that emissions causing climate change are to remain exactly as they are now, neither increasing nor decreasing).
+This limits further the time frame for which projections and models are viable.
+Finally, the resolution of our data is not ideal; we have monthly aggregate data so we are projecting out monthly as well.
+This is not as helpful as a finer resolution may have been given the fact that monthy aggregate data does not accurately reflect the extreme temeprature spikes we may see on specific days.
+In the future, it would be usely to repeat this modeling ideally with daily data.
+Even geting multiple temepratures a day could had more complexity as we should except to see patterns of temeprature changes throughout the day.
+
+**Selected Refrences**
+Johnson, Z.
+I., Wheeler, B.
+J., Blinebry, S.
+K., Carlson, C.
+M., Ward, C., & Hunt, D.
+E.
+(2013a).
+Dramatic variability of the carbonate system at a temperate coastal ocean site (Beaufort, North Carolina, USA) is regulated by physical and biogeochemical processes on multiple timescales.
+PloS One, 8(12), e85117.
+https://doi.org/10.1371/journal.pone.0085117
+
+McIntyre, A. D. (2010). Life in the world’s oceans. In Wiley eBooks. https://doi.org/10.1002/9781444325508
+
+Wang, Z., Tsementzi, D., Williams, T.
+C., Juarez, D.
+L., Blinebry, S.
+K., Garcia, N.
+S., Sienkiewicz, B.
+K., Konstantinidis, K.
+T., Johnson, Z.
+I., & Hunt, D.
+E.
+(2020a).
+Environmental stability impacts the differential sensitivity of marine microbiomes to increases in temperature and acidity.
+The ISME Journal, 15(1), 19–28.
+https://doi.org/10.1038/s41396-020-00748-2

### 2024-04-23 Draft_report.Rmd
YilunZhu did conclusion part

@@
 ```{r, echo=FALSE}
-#The ACF and Pacf tell us that there is ovbious and strong seasonality and some auto-regressive component up to about 6 lags.
+#The ACF and Pacf tell us that there is obvious and strong seasonality and some auto-regressive component up to about 6 lags.
 P1 <-Acf(temperature_ts, main = "ACF on original data")
@@
 # Model 5:  SSES 
-SSES_scores <- accuracy(NN_for$mean, test_ts)
+SSES_scores <- accuracy(SSES_seas$forecast, test_ts)
 
 # Model 6:  Neural Network 
-NN_scores <- accuracy(SSES_seas$forecast, test_ts)
+NN_scores <- accuracy(NN_for$mean, test_ts)
 
@@
 The first is regarding our prediction time frame We only have access to ~ 10 years of data which likely does not demonstrate enough of a trend to be recognized as the role of climate change in our models.
-We should expect increasing temepratures in the future as a result of climate change.
-Even if our model did capture this linear trend, the modeling would be based on the assumtions that the trend is constant (aka that emissions causing climate change are to remain exactly as they are now, neither increasing nor decreasing).
+We should expect increasing temperatures in the future as a result of climate change.
+Even if our model did capture this linear trend, the modeling would be based on the assumptions that the trend is constant (aka that emissions causing climate change are to remain exactly as they are now, neither increasing nor decreasing).
 This limits further the time frame for which projections and models are viable.
 Finally, the resolution of our data is not ideal; we have monthly aggregate data so we are projecting out monthly as well.
-This is not as helpful as a finer resolution may have been given the fact that monthy aggregate data does not accurately reflect the extreme temeprature spikes we may see on specific days.
+This is not as helpful as a finer resolution may have been given the fact that monthly aggregate data does not accurately reflect the extreme temperature spikes we may see on specific days.
 In the future, it would be usely to repeat this modeling ideally with daily data.
-Even geting multiple temepratures a day could had more complexity as we should except to see patterns of temeprature changes throughout the day.
+Even getting multiple temperatures a day could had more complexity as we should except to see patterns of temperature changes throughout the day.
 
-**Selected Refrences**
+**Conclusion**
+By modeling seawater surface temperature using six models and initially decomposing the data itself, we found that at least the ambient seawater surface temperature obtained from Duke Marine Lab did not show a clear upward trend during this decade.
+The conclusion seems somewhat inconsistent with the current impact of global climate change on seawater temperatures.
+However, this trend should be independent of our processing of the data, that is, averaging the weekly data into monthly data because the overall water temperature changes should not be diluted by just taking the monthly average.
+So, there may be something worth studying and discussing behind it.
+After comparing the errors of the six models with the original data, we finally found that the SSES model had the smallest error, which can provide a reference for subsequent researchers to choose water temperature modeling.
+Since we did not use other relevant variables as predictors, we believe that the current water temperature model still has much room for improvement.
+
+**Selected Refernces**
 Johnson, Z.
